---
title: "Reproduce.Rmd"
output: html_document
---
Load the required libraries
```{r echo=FALSE}
library(dplyr)
library(ggplot2)

```
Load the data from the file into mydf variable
the file should be located in the default directory of R

```{r}
setwd("E:/r/reproduce/")
data_row <- read.csv("activity.csv")
```
Remove NA from the dataset
```{r}
data <- data_row[ with (data_row, { !(is.na(steps)) } ), ]

```

Calculating the mean total number of steps taken per day
- group the data by date and put the result in new variable
- use summarise function to calculate the sum of the steps per day

```{r}
by_day <- group_by(data, date)
steps_by_day <- summarise(by_day, total = sum(steps))
steps_by_day

```

Make a histogram of the steps taken by day

```{r}
hist(steps_by_day$total, main="Histogram of total number of steps per day", 
     xlab="Total number of steps in a day")
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
v_mean <- (mean(steps_by_day$steps))
v_median <- factor(median(steps_by_day$steps))


```
Printing the mean
```{r}
v_mean
```
Printing the median
```{r}
v_median
```
Calculate the average daily activity pattern
```{r}
by_interval <- group_by(data, interval)
steps_by_interval <- summarise(by_interval, steps = mean(steps))
```
Create a time series plot 
```{r}
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

- find row with max of steps
- find interval with this max
```{r}
max_steps_row <- which.max(steps_by_interval$steps)
steps_by_interval[max_steps_row, ]
```

Imputing missing values
Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data_row))
```

Replacing NA's with the mean for that 5-minute interval.
```{r}
data_imputed <- data_row
for (i in 1:nrow(data_imputed)) {
  if (is.na(data_imputed$steps[i])) {
    interval_value <- data_imputed$interval[i]
    steps_value <- steps_by_interval[
      steps_by_interval$interval == interval_value,]
    data_imputed$steps[i] <- steps_value$steps
  }
}
```

Create new dataset df_imputed_steps_by_day which equals to data_row but without NA's. All NA's are replaced with mean of 5-minute interval.
```{r}
df_imputed_steps_by_day <- aggregate(steps ~ date, data_imputed, sum)
head(df_imputed_steps_by_day)
```
Drawing the Histogram of total number of steps per day for the new Dataset 
```{r}
hist(df_imputed_steps_by_day$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps in a day")
```

Get mean and median of imputed data
1- mean
```{r}
mean(df_imputed_steps_by_day$steps)
```
2- median 
```{r}
median(df_imputed_steps_by_day$steps)
```

Get mean and median of original data
1- mean
```{r}
mean(steps_by_day$total)
```
2- median 
```{r}
median(steps_by_day$total)
```
Mean value stays the same but there is slight difference in meadian value.


Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
data_imputed['type_of_day'] <- weekdays(as.Date(data_imputed$date))
data_imputed$type_of_day[data_imputed$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
data_imputed$type_of_day[data_imputed$type_of_day != "weekend"] <- "weekday"
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
# convert type_of_day from character to factor
data_imputed$type_of_day <- as.factor(data_imputed$type_of_day)

# calculate average steps by interval across all days
df_imputed_steps_by_interval <- aggregate(steps ~ interval + type_of_day, data_imputed, mean)

# creat a plot
qplot(interval, 
      steps, 
      data = df_imputed_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
  facet_wrap(~ type_of_day, ncol = 1)
```